---
title: "Still Suburban"
author:
- Nathan Lauster
- Dmitry Shkolnik
- Jens von Bergmann
date: '2018-08-21'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE,
	fig.width = 9
)
library(tidyverse)
library(cancensus)
library(cancensusHelpers)
library(tongfen)
library(sf)
```

```{r}
commute_vars_2016 <- search_census_vectors("Main mode of commuting","CA16","Total") %>% 
  bind_rows(child_census_vectors(.))
commute_vars_2006 <- search_census_vectors("mode of","CA06","Total") %>% 
  bind_rows(child_census_vectors(.))
all_vectors <- c(commute_vars_2006$vector,commute_vars_2016$vector)
cmas <- list_census_regions("CA16") %>% filter(level=="CMA")

canada_data <- get_census("CA16",regions=list(C="01"),vectors = all_vectors,level="Regions",labels ="short") %>%
  mutate(`Active 2006`=(v_CA06_1104+v_CA06_1105)/v_CA06_1100,
           `Transit 2006`=(v_CA06_1103)/v_CA06_1100,
           `Active 2016`=(v_CA16_5804+v_CA16_5807)/v_CA16_5792,
           `Transit 2016`=(v_CA16_5801)/v_CA16_5792)

type_colors <- c("Unclassified"="#808080","Exurban"="#ffffff","Auto Suburb"="#FDFFB0","Transit Suburb"="#E0E507","Active Core"="#989A07")

tongfen_data_for_region <- function(region){
  data <- get_tongfen_census_ct(region %>% as_census_region_list,
                                vectors=all_vectors,geo_format='sf') %>%
    mutate(`Density 2006`=Population_CA06/`Shape Area`,
           `Density 2016`=Population_CA16/`Shape Area`,
           `Active 2006`=(v_CA06_1104+v_CA06_1105)/v_CA06_1100,
           `Transit 2006`=(v_CA06_1103)/v_CA06_1100,
           `Active 2016`=(v_CA16_5804+v_CA16_5807)/v_CA16_5792,
           `Transit 2016`=(v_CA16_5801)/v_CA16_5792
           ) %>%
    left_join((.) %>% 
                st_set_geometry(NULL) %>%
                select(all_vectors,"Population_CA06","Population_CA16","Shape Area","CMA_UID") %>%
                group_by(CMA_UID) %>%
                summarize_all(sum,na.rm=TRUE) %>%
                set_names(paste0("CMA_",names(.))),by=c("CMA_UID"="CMA_CMA_UID")) %>%
    mutate(`CMA_Density 2006`=CMA_Population_CA06/`CMA_Shape Area`,
           `CMA_Density 2016`=CMA_Population_CA16/`CMA_Shape Area`,
           `CMA_Active 2006`=(CMA_v_CA06_1104+CMA_v_CA06_1105)/CMA_v_CA06_1100,
           `CMA_Transit 2006`=(CMA_v_CA06_1103)/CMA_v_CA06_1100,
           `CMA_Active 2016`=(CMA_v_CA16_5804+CMA_v_CA16_5807)/CMA_v_CA16_5792,
           `CMA_Transit 2016`=(CMA_v_CA16_5801)/CMA_v_CA16_5792) %>%
    mutate(Type=case_when(
      `Density 2016`<150 ~ "Exurban",
      is.na(`Active 2016`) | is.na(`Transit 2016`) ~ "Unclassified",
       `Active 2016` < 1.5* `CMA_Active 2016` & `Transit 2016` < 1.5* `CMA_Transit 2016` ~ "Auto Suburb",
      `Active 2016` < 1.5* `CMA_Active 2016` & `Transit 2016` >= 1.5* `CMA_Transit 2016` & `Transit 2016` > 0.5 * canada_data$`Transit 2016` ~ "Transit Suburb",
      `Active 2016` > 0.5 * canada_data$`Active 2016` ~ "Active Core",
      TRUE ~ "Unclassified"
    )) %>%
    mutate(Type=factor(Type,levels=names(type_colors)))
  data
}


data_for_region <- function(region,geo_level="CT"){
  vectors <- commute_vars_2016$vector
  data <- get_census("CA16",
                     region %>% as_census_region_list,
                     level=geo_level,
                     label="short",
                     vectors=vectors,
                     geo_format='sf') %>%
    mutate(`Density`=Population/`Shape Area`,
           `Active`=(v_CA16_5804+v_CA16_5807)/v_CA16_5792,
           `Transit`=(v_CA16_5801)/v_CA16_5792
           ) %>%
    left_join((.) %>% 
                st_set_geometry(NULL) %>%
                select(vectors,"Population","Shape Area","CMA_UID") %>%
                group_by(CMA_UID) %>%
                summarize_all(sum,na.rm=TRUE) %>%
                set_names(paste0("CMA_",names(.))),by=c("CMA_UID"="CMA_CMA_UID")) %>%
    mutate(`CMA_Density`=CMA_Population/`CMA_Shape Area`,
           `CMA_Active`=(CMA_v_CA16_5804+CMA_v_CA16_5807)/CMA_v_CA16_5792,
           `CMA_Transit`=(CMA_v_CA16_5801)/CMA_v_CA16_5792) %>%
    mutate(Type=case_when(
      `Density`<150 ~ "Exurban",
      is.na(`Active`) | is.na(`Transit`) ~ "Unclassified",
       `Active` < 1.5* `CMA_Active` & `Transit` < 1.5* `CMA_Transit` ~ "Auto Suburb",
      `Active` < 1.5* `CMA_Active` & `Transit` >= 1.5* `CMA_Transit` & `Transit` > 0.5 * canada_data$`Transit 2016` ~ "Transit Suburb",
      `Active` > 0.5 * canada_data$`Active 2016` ~ "Active Core",
      TRUE ~ "Unclassified"
    )) %>%
    mutate(Type=factor(Type,levels=names(type_colors)))
  data
}
```

## Questions
I have a couple of questions regarding the classification. The categories in the [working paper](http://www.canadianurbanism.ca/wp-content/uploads/2018/08/Still-Suburban-report-For-Release-comp.pdf) aren't mutually exclusive, leaving "unclassified" areas. One such category is a tract with transit use > 150% of CMA but less than 50% of national average that is not an urban core. Similarly, a tract with active transport > 150% of CMA average but less than 50% of national average is unclassified (even if transit use is high). Lastly, I don't understand the definition of "Exurb", in particular the exact meaning of "more than 50% of workers commuting into the metropolitan area".

## Geographies
We can chose to map the data on 2016 geography, or on a common tiling for 2006 and 2016 census tracts using the [tongfen package](https://github.com/mountainMath/tongfen). If we use data from a single census we have the choice between using CT or DA level data.


## Vancouver region
```{r}
data <- data_for_region(cmas %>% filter(grepl("Vancouver",name)),"CT")

ggplot(data,aes(fill=Type)) +
  geom_sf(size=0.1) +
  theme_void() +
  scale_fill_manual(values=type_colors) +
  coord_sf(datum=NA)
```

## Toronto
```{r}
data <- data_for_region(cmas %>% filter(grepl("Toronto",name)))

ggplot(data,aes(fill=Type)) +
  geom_sf(size=0.1) +
  theme_void() +
  scale_fill_manual(values=type_colors) +
  coord_sf(datum=NA)
```

## Montreal
```{r}
data <- data_for_region(cmas %>% filter(grepl("Montr",name)))

ggplot(data,aes(fill=Type)) +
  geom_sf(size=0.1) +
  theme_void() +
  scale_fill_manual(values=type_colors) +
  coord_sf(datum=NA)
```

## Calgary
```{r}
data <- data_for_region(cmas %>% filter(grepl("Calgary",name)))

ggplot(data,aes(fill=Type)) +
  geom_sf(size=0.1) +
  theme_void() +
  scale_fill_manual(values=type_colors) +
  coord_sf(datum=NA)
```


## Vancouver DA level

```{r}
data <- data_for_region(cmas %>% filter(grepl("Vancouver",name)),"DA")

ggplot(data,aes(fill=Type)) +
  geom_sf(size=0.1) +
  theme_void() +
  scale_fill_manual(values=type_colors) +
  coord_sf(datum=NA)
```


## K-menas clustering
We cluster based on transit mode share, active transport mode share and density.
```{r}
library(broom)

kd <- data %>% 
  mutate(Density=log(Density+1),
         Active=log(Active+1),
         Transit=log(Transit+1)) %>%
  mutate(Density=Density/max(Density,na.rm=TRUE),
         Active=Active/max(Active,na.rm=TRUE),
         Transit=Transit/max(Transit,na.rm=TRUE)) %>%
  select(Density,Active,Transit) %>%
  na.omit()

clusters <- kmeans(kd %>%   st_set_geometry(NULL),4 )

kd$cluster <- as.factor(clusters$cluster)

ggplot(kd, aes(fill=cluster)) +
  geom_sf() +
  coord_sf(datum=NA)

```

### Projections
```{r}
ggplot(kd,aes(x=Density,y=Transit,color=cluster)) +
  geom_point()
```

```{r}
ggplot(kd,aes(x=Density,y=Active,color=cluster)) +
  geom_point()
```


```{r}
ggplot(kd,aes(x=Active,y=Transit,color=cluster)) +
  geom_point()
```